\chapter{Datasets and running conditions}
\label{chap:prod:data}

The data used for this measurement were collected at the very beginning of 
\runtwo\ of the \ac{LHC}, during a special 15-day `early measurements' period.
The number of bunches in the accelerator was gradually increased from 50 to 482 
bunches at the end of the period, with the number colliding at IP8 increasing 
from 36 to 397 bunches.
In addition, the minimum bunch spacing was set to \SI{50}{\nano\second}, as in 
\runone, rather than the nominal \runtwo spacing of \SI{25}{\nano\second}.
These steps were taken primarily for machine safety, as the \ac{LHC} began 
operating at a new centre-of-mass energy, \sqrtseq{13}, never before seen at a 
particle collider.

The machine suffered several technical problems during this period, and so a 
significant amount of time allocated for `physics running' was lost to 
recovery.
At \lhcb, each \ac{LHC} fill is indexed with a number, and this measurement 
uses fills 3976, 3981, 3988, 3992, and 3996.
The combination of these fills corresponds to around \SI{32}{\hour} of 
collisions.
Other fills taken during the early measurements period were excluded from the 
analysis due to them either having short times, usually corresponding to 
unstable beams unfit for physics, or because one or more sub-detector was found 
to be operating incorrectly during data-taking.

The combined dataset used for the measurement corresponds to an integrated 
luminosity, defined in \cref{eqn:prod:introduction:integrated_lumi}, of
\begin{equation}
  \intlumi = \SI{\xsectotlumi}{\per\pico\barn}.
  \label{eqn:prod:xsectotlumi}
\end{equation}
For reference, the total integrated luminosity collected in \runone\ at \lhcb\ 
was \SI{3}{\per\femto\barn}, 600 times more, whilst that for the \sqrtseq{7}\ 
\lhcb\ open charm production measurement was \SI{15}{\per\nano\barn}, 300 times 
less.
All data were taken with \lhcb\ dipole magnet in the `down' polarity, and were 
processed via the Turbo stream data flow, as described in 
\cref{chap:intro:lhcb:detector}.
The selection of events and charm candidates, both in the trigger and offline, 
will be described in \cref{chap:prod:sel}.
A description of the luminosity measurement will now be given.

\section{Luminosity determination}
\label{chap:prod:data:lumi}

The luminosity was measured in a dedicated analysis as part of the general 
\lhcb\ early measurements campaign.
The luminosity was used as input not only for the open charm production 
analysis discussed here, but also in the measurements of \PJpsi and \PB hadron 
production~\cite{LHCb-PAPER-2015-037,LHCb-PAPER-2016-031}, which were also part 
of that campaign.

The instantaneous luminosity is defined in \cref{eqn:intro:lhc:inst_lumi}, and 
can be re-formulated more succinctly as
\begin{equation}
  \lumi = N_{\text{p}}^{2}\revfreq\Omega,
  \label{eqn:prod:data:lumi}
\end{equation}
where $\Omega$ encapsulates the overlapping bunch densities.
If each bunch, 1 and 2, has particle density distributions $\rho_{1}(x, y, z, 
t)$ and $\rho_{2}(x, y, z, t)$, the overlap can be expressed as
\begin{equation}
  \Omega \propto \int \rho_{1}(x, y, z, t)\rho_{2}(x, y, z, t)
                      \dif{x}\dif{y}\dif{z}\dif{t}.
  \label{eqn:prod:data:lumi:profile}
\end{equation}
At \lhcb, two methods have been used to measure the bunch profiles.
% TODO ok, but what is done with the scan data?
The first is the \ac{VDM} scan, where the beams are scanned across each other 
in fine $x$ and $y$ steps to measure the transverse profile, which is also used 
in the \atlas, \cms, and \alice\ experiments.
It assumes that the profiles in $x$ and $y$ are factorisable, such as being two 
independent Gaussian distributions.
The high precision of the \lhcb\ \velo\ sub-detector allows for the second 
method: the \ac{BGI} scan.
% TODO describe BGI principles
This method

As the \ac{VDM} and \ac{BGI} methods are independent, there serve as 
cross-checks of one another and can be combined in an average if they are 
compatible.
They were both used in an analysis to measure the \runone\ integrated 
luminosity, where a relative uncertainty of \SI{1.16}{\percent} was achieved, 
the highest precision at any \ac{LHC} experiment~\cite{LHCb-PAPER-2014-047}.
For the early measurements campaign, only the data for the \ac{BGI} measurement 
was used

% Why does he BGI require special conditions?
Both the \ac{VDM} and the \ac{BGI} methods require special \ac{LHC} running 
conditions, the data collected during which is not used for physics analyses.
The absolute luminosity is then determined for some specific fills, and it is 
then necessary to somehow translate this to a luminosity measurement for the 
fills used for physics data.
% TODO describe how the relative measurements are performed
