\chapter{Acknowledgements}

So many people have done so much for me over the past four-or-so years, and 
this thesis would never have been finished without them.
To anyone who I've had even the most minimal contact with, but I fail to 
mention: you have truly helped me, and I am very grateful for it.
Still, I must be particular.

My supervisors George Lafferty and Sajan Easo helped me immensely when I was 
starting; Sajan during my Master's project, and George during my PhD.
Patrick Spradlin was my third supervisor in everything but name, and did a huge 
amount to help.
Each of them handled my endless questions with patience, and taught me 
everything I wanted to know.

To everyone I shared an office with at Manchester, Lorenzo, Shanzhen, Giulio, 
Suzanne, and Kevin, thanks for making being sat in front of a computer all day 
a bit more bearable.
Particular thanks to Igor, for showing me the world outside, and Jon, for 
telling what were only barely passable as jokes.

Thanks to Chris and Dominik, for sharing the cross-section experience, putting 
up with the hard times and silly hours, and finishing with an amazing piece of 
work.
I'm really proud of what we managed to do, and I couldn't have done it with 
anybody else; it's really more your achievement than mine.

I must thank Chris and Marco for allowing me to study at Manchester and for the 
invaluable guidance, and the whole the Manchester group for being nice people 
to be around.
Also the University, the Rutherford Appleton Laboratory, and the Science \& 
Technology Facilities Council for funding this whole business, and to Fred for 
so deftly pulling the purse strings to allow it to happen.

Vava Gligorov and Patrick Koppenburg have done a lot more for more than they 
can know, setting me down a path I couldn't have gotten on without them.
I'm grateful for their wisdom, guidance, and optimism.

Finally, thanks to everyone within the \lhcb\ collaboration.
It's a fantastic environment to work in, and the people I've worked with and 
spoken to have taught me so much. Coordinators, conveners, reviewers, 
colleagues, people I only know on mailing lists, and friends, I owe you all.
