\chapter{Summary}
\label{chap:cpv:summary}

The violation of the \CP\ symmetry has not yet been observed in charm decays, 
the last system in the \ac{SM} where it is predicted.
Measuring \CP\ asymmetries can be experimentally challenging as a good 
understanding of production and detection asymmetries must be obtained, but the 
construction of the \CP\ asymmetry difference \dACP\ allows for robust 
measurements that are less sensitive to spurious background effects.

With the analysis presented in this \namecref{chap:cpv}, the first measurement 
of direct \CP\ violation in charm baryons at \lhcb\ has been completed.
The decays \LcTopKK\ and \LcToppipi\ are reconstructed via the decay 
\LbToLcmuX, with the \PLambdab\ and muon providing powerful background 
discrimination.
Backgrounds from charm mesons and removed with mass vetoes.
Contributions from background asymmetries from the \PLambdab\ production 
asymmetry and the pion, kaon, proton, and muon detection asymmetries are 
suppressed by a weighting between the \PLambdac\ decay modes, employing a 
\acl{BDT} to partition the data in \PLambdac\ and proton kinematics.
The acceptance and selection efficiencies are computed as a function of the 
five-dimensional \PLambdac\ phase space also using a \ac{BDT}, with an average 
efficiency of around \SI{1}{\percent}.
The \PLambdac\ signal yields are counted with \chisq\ fits, weighting the data 
with the kinematic weights and the inverse of the phase space efficiencies.
Systematic effects due to the signal and background modelling in the fit, the 
residual background asymmetries after the weighting, and the contribution of 
\PLambdac\ candidates from sources other than semileptonic \PLambdac\ decays 
are studied, and are found to be considerably smaller than the statistical 
uncertainty on the signal yields.

The difference in \CP\ asymmetries in the decays \LcTopKK\ and \LcToppipi is 
measured to be
\begin{align*}
  \dACP &= \ACP(\pKK) - \ACP(\ppipi) \\
        &= \SIerrs{\deltaACPval}{\deltaACPunc}{0.28}{\percent}.
\end{align*}
The central value is currently blinded, pending approval from the \lhcb\ 
collaboration.

Measurements of \dACP\ using \PDzero decays at \lhcb\ spurred a large interest 
from the theory community, creating much-needed effort to understand the 
mechanisms of \CP\ violation in hadronic charm decays.
Evidence of \CP\ violation in charm baryon decays is likely to spark similar 
levels of curiosity.
If it is not found in this analysis, it is possible that \lhcb\ may find it 
elsewhere in the near future, as it continues to grow the worlds largest 
samples of charm decays.
